Data Migration under a Changing Schema

### 2023-07-06 2022Migration/articleMigration.tex

   Information systems are typically used by actors (both users and computers) who are distributed and work with data all the time.
   As a consequence, the data in a system changes continually.
   In practice, actors ``talk to'' the system through an ingress mechanism, which connects each user to the right service(s).
   The ingress function is provided by the deployment platform, so it is beyond the scope of this paper.
   Also, every service runs independent of other services,
   meaning that each service can be stopped, (re)started and substituted without disrupting the other services.
   For practical purposes, this presumes a platform of managed services, such as Kubernetes,
   even though the theory in this paper is not specific to any platform.

@@

   \includegraphics[scale=0.1]{figures/Before migration.png}

@@

\item From schemas $\schema$ and $\schema'$, a migration engineer%

@@

      generates a migration script, using the Ampersand compiler.
      This script specifies the migration system, $\migrsys$, that includes $\infsys$, $\infsys'$ and a number of migration rules.
      $\migrsys$ is meant to migrate the data and ensure that $\dataset'$ satisfies all invariants of $\schema'$.

@@

      In specific cases, the schema $\schema_{\migrsys}$ will be enough to perform the migration automatically.
      As soon as the data is transferred and $\migrsys$ has satisfied the invariants of $\schema'$,
      $\infsys'$ can take care of production on its own without the help of $\migrsys$.

@@

      For that purpose, she may want to alter the migration script to accommodate such requirements.

@@

      But she can also ask business users to resolve issues by specifying process rules.

@@

      She can test the resulting migration script, $\migrsys'$ separately before taking the desired system to production.

@@

      Then, the migration engineer deploys system $\migrsys'$.

@@

   \includegraphics[scale=0.1]{figures/4MoD.png}

@@

      Let us call this moment the moment of deployment (MoD).

@@

      During this phase, the ingress is still referring users to the existing system,

@@

      During this phase, users can still change $\dataset$ (both inserts and deletes) while $\infsys$ is keeping its invariants true.

@@

@@

   \includegraphics[scale=0.1]{figures/5MoT.png}

@@

>>>>

   Information systems are typically used by actors (both users and computers) who are distributed and work with data all the time.
   As a consequence, the data in a system changes continually.
   In practice, actors ``talk to'' the system through an ingress mechanism,
   which routes the traffic between each user and the required application service(s).
   The ingress function is used during migration to switch users from the old to new functionality.
   In this research we assume that the deployment platform provides the ingress mechanism,
   so developers merely have to configure it.
   Also, we assume that every service runs independent of other services,
   meaning that each service can be stopped, (re)started and substituted without disrupting the other services.
   For practical purposes, this presumes a platform of managed services, such as Kubernetes,
   even though the theory in this paper is not specific to any platform.

@@

   \includegraphics[scale=0.09]{figures/Before migration.png}

@@

\item A migration engineer%

@@

      uses the Ampersand compiler to generate a migration script from schemas $\schema$ and $\schema'$.
      This script specifies a migration system, $\migrsys$, that includes $\infsys$, $\infsys'$ and a number of migration rules.
      $\migrsys$ is meant to migrate the data and ensure that $\dataset'$ satisfies all invariants of $\schema'$
      before taking $\infsys'$ to production.

@@

      In specific cases, $\migrsys$ suffices to perform the migration automatically.

@@

      For that purpose, she may want to alter the migration script to accommodate such requirements, yielding another migration script $\migrsys'$.

@@

      But she can also choose to specify process rules that require business users to resolve issues interactively.

@@

      She can test $\migrsys'$ separately before taking the desired system to production.

@@

      At this moment, let us call it the moment of deployment (MoD),
\begin{figure}[h]

@@

   \includegraphics[scale=0.09]{figures/4MoD.png}

@@

      the migration engineer deploys system $\migrsys'$.

@@

      During this phase, the ingress still refers users to the existing system,

@@

      Users can still change data in $\dataset$, while $\infsys$ is keeping its invariants true.

@@

\begin{figure}[h]

@@

   \includegraphics[scale=0.09]{figures/5MoT.png}

@@

      Simultaneously, all clients must switch to the functionality of $\infsys'$.
